---
title: "Practical ML - Course project"
author: "Andrea Eoli [20-02-2023]"
#date: "`r Sys.Date()`"
output: html_document
---

\fontsize{10}{12}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width='90%')
library(ggplot2)
library(dplyr)
library(caret)
library(parallel) # running caret is too computationally intense and slow without parallel
library(doParallel)

```

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (section 'Weight Lifting Exercise Dataset').


## Task
The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.


## Analysis results 
After selecting the relevant columns (predictors) from the training data set, data gets preprocessed by removing columns with too many NAs (threshold 95%) and columns with near zero values. The final training data set has 53 columns. It gets partitioned into 70% training data and 30% testing data. The training data gets processed with 3 models, whose accuracy will be compared: Random Forest, Stochastic Gradient Boosting, and Linear Discriminant Analysis. Their accuracy is, respectively, 99.34%, 96.14%, and 70.46%. Therefore, Random Forest is the model with the highest accuracy and the smallest out-of-sample error (0.7%), and is used to predict the "classe" of the 20 cases (validation data).


```{r data_load_and_prep}
training <- read.csv("pml-training.csv", header = TRUE)
validation <- read.csv("pml-testing.csv", header = TRUE)

# Find relevant columns
predictors <- grepl("belt|[^(fore)]arm|dumbbell|forearm|classe", names(training)) 
train_filt <- training %>% select(names(training[,predictors])) 

# Remove NAs
NA_columns <- apply(train_filt, 2, function(x) sum(is.na(x)) > dim(train_filt)[1]*0.95  || sum(x=="") > dim(train_filt)[1]*0.95)
train_filt <- train_filt[,!NA_columns]

# Filter by NZV
NZV_cols <- nearZeroVar(train_filt, saveMetrics = TRUE)
train_filt <- train_filt[, NZV_cols$nzv==FALSE]

# Convert class and select same cols from validation dataset
train_filt$classe <- as.factor(train_filt$classe) # label variable
valid_filt <- validation %>% mutate(classe = NA) %>% select(names(train_filt)) 

# Prepare partition and DFs
inTrain <- createDataPartition(train_filt$classe, p = 0.7)[[1]]
train <- train_filt[inTrain,]
test <- train_filt[-inTrain,]

```

```{r model}
cluster <- makeCluster(detectCores() - 1) # convention to leave 1 core for OS
registerDoParallel(cluster) # without this it takes more time than the last movie of Park Chan-wook 

fitControl <- trainControl(method = "cv", # We use cross-validation as resampling method
                           number = 5, # and it's going to be a 5-fold CV
                           allowParallel = TRUE) # this tells caret to use the 'parallel' package

mod_rf <- train(classe ~ ., data=train, method="rf", trControl = fitControl)
mod_gbm <- train(classe ~ ., data=train, method="gbm", trControl = trainControl(allowParallel = T))
mod_lda <- train(classe ~ ., data=train, method="lda")
stopCluster(cluster) # to explicitly shut down the cluster

```

```{r results}
test_rf <- predict(mod_rf, test)
test_gbm <- predict(mod_gbm, test)
test_lda <- predict(mod_lda, test)

confusionMatrix(test_rf, test$classe) # Accuracy: 0.9934 --> best model
confusionMatrix(test_gbm, test$classe)$overall[1] # Accuracy: 0.9614
confusionMatrix(test_lda, test$classe)$overall[1] # Accuracy: 0.7046
```

```{r prediction}
predict(mod_rf, valid_filt)

```



