---
title: 'Final Project: Build a Shiny app'
author: "Andrea Eoli"
date: "06 May 2023"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Instructions

1. Write a shiny app with documentation.
2. Deploy it on Rstudio's shiny server
3. Share the app link
4. Share the server.R and ui.R scripts on github

## How to use the application

Start to write a short sentence and the app will predict which words are most likely to follow.

The application is running here: https://eolia.shinyapps.io/Week4_AE/

ui.R, and Server.R Code in my github repository 

## Galton Families Dataset

The data used in the app comes from the GaltonFamilies dataset. 

```{r summary}
library(HistData)
data(GaltonFamilies)
summary(GaltonFamilies)
```

## Plot
In the scatterplot below are represented the data used for the prediction model
```{r plot}
library(HistData)
data(GaltonFamilies)
plot(jitter(GaltonFamilies$childHeight) ~ GaltonFamilies$midparentHeight,xlab="Average Height of the Parents (in inches)",ylab="Height of the Child (in inches)",main="Scatterplot of Galton Family Data",pch=19,frame.plot=FALSE,col=ifelse(GaltonFamilies$gender=="female", "pink", "light blue"))
legend(65,80,pch=c(19,19),col=c("pink","light blue"),c("female", "male"),bty="o",cex=.8)
```
