---
title: 'Final Project: Capstone'
author: "Andrea Eoli"
date: "15 May 2023"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Instructions

1. Write a shiny app with documentation.
2. Deploy it on Rstudio's shiny server
3. Share the app link
4. Share the server.R and ui.R scripts on github


**Course**: "Data Science Capstone"

**Specialization**: "Data Science"

## How to use the application

The app is a word predictor. Simply start to write a short sentence and the app will predict which words are most likely to follow.

[Click here to access the app](https://eolia.shinyapps.io/Week4_AE/)


## Data sources

The data comes from 2 sources:  
1. *Ngrams* is a dataframe with words ('words') that follow an input string ('ngrams') and their frequency of appearance ('freq')
2. *Contractions* contains the info to convert contracted words into single ones (e.g., "it's" to "it is")

```{r summary}
load(file="ngrams")
subset <- which(stringr::str_count(ngrams$ngrams) > 8)
head(ngrams[subset,])
```

## Code and GitHub
The ui.R and Server.R scripts are available [in my github repository](https://github.com/aeoli/DataScience_JHU_Coursera/tree/master/9.%20Final%20capstone/Final)
